---
title: "Présentation de l'application de colSBM sur Doré et al. 2020"
output:
    html_document:
        toc: true
        theme: united
---

```{r require_libn, echo = FALSE, include=FALSE}
require("ggplot2")
require("dplyr")
require("knitr")
devtools::load_all("R/")
source("wip/temporary_plot.R")
```

```{r pretty_matrix_print, echo = FALSE}
# Define a generic method that transforms an object x in a LaTeX string
as_latex <- function(x, ...) {
  UseMethod("as_latex", x)
}

# Define a class latex for LaTeX expressions
as_latex.character <- function(x) {
  structure(
    paste(x, collapse = " "),
    class = c("latex", "character")
  )
}

# A character string of class latex is rendered in display mode
# Define a knit_print() method for the latex class
knit_print.latex <- function(x, ...) {
  knitr::asis_output(
    paste0("$$", x, "$$")
  )
}

# Now, define a method as_latex for matrix
as_latex.matrix <- function(x, ...) {
  as_latex(c(
    "\\begin{bmatrix}",
    paste(
      t(x),
      rep(c(rep("&", nrow(x) - 1), "\\\\"), ncol(x)),
      collapse = ""
    ),
    "\\end{bmatrix}"
  ))
}

# Indicate to knitr that matrix are rendered as latex
knit_print.matrix <- function(x, ...) {
  knitr::knit_print(as_latex(round(x, 2)))
}
```

```{r taxonomy_functions, echo = FALSE}
interaction_data <- read.table(file = "real_data/data/interaction-data.txt", sep = "\t", header = TRUE)

insect_orders <- unique(interaction_data$insectorder)
plant_family <- unique(interaction_data$plantfamily)
### Matching taxonomy
taxonomy_in_clusters <- function(unlisted_model) {
  if (is.list(unlisted_model)) {
    lapply(seq_len(length(unlisted_model)), function(col_idx) {
      # Per collection
      # Empty init
      insect_count <- t(sapply(insect_orders, function(order) {
        out_count <- rep(0, unlisted_model[[col_idx]]$Q[2])
        out_count
      }))

      plant_count <- t(sapply(plant_family, function(order) {
        out_count <- rep(0, unlisted_model[[col_idx]]$Q[1])
        out_count
      }))

      for (m in seq.int(unlisted_model[[col_idx]]$M)) {
        #### Insect
        insect_names <- names(unlisted_model[[col_idx]]$Z[[1]][[2]])

        insect_count <- insect_count + t(sapply(insect_orders, function(order) {
          out_count <- rep(0, unlisted_model[[col_idx]]$Q[2])
          names(out_count) <- seq.int(unlisted_model[[col_idx]]$Q[2])
          insect_count <- table(unlisted_model[[col_idx]]$Z[[m]][[2]][grep(order, insect_names)])
          out_count[names(insect_count)] <- insect_count
          out_count
        }))
        #### Plants
        plant_names <- names(unlisted_model[[col_idx]]$Z[[1]][[1]])

        plant_count <- t(sapply(plant_family, function(order) {
          out_count <- rep(0, unlisted_model[[col_idx]]$Q[1])
          names(out_count) <- seq.int(unlisted_model[[col_idx]]$Q[1])
          plant_count <- table(unlisted_model[[col_idx]]$Z[[m]][[1]][grep(order, plant_names)])
          out_count[names(plant_count)] <- plant_count
          out_count
        }))
      }
      return(list(insects = insect_count, plants = plant_count))
    })
  } else {
    # Per collection
    # Empty init
    insect_count <- t(sapply(insect_orders, function(order) {
      out_count <- rep(0, unlisted_model$Q[2])
      out_count
    }))

    plant_count <- t(sapply(plant_family, function(order) {
      out_count <- rep(0, unlisted_model$Q[1])
      out_count
    }))

    for (m in seq.int(unlisted_model$M)) {
      #### Insect
      insect_names <- names(unlisted_model$Z[[1]][[2]])

      insect_count <- insect_count + t(sapply(insect_orders, function(order) {
        out_count <- rep(0, unlisted_model$Q[2])
        names(out_count) <- seq.int(unlisted_model$Q[2])
        insect_count <- table(unlisted_model$Z[[m]][[2]][grep(order, insect_names)])
        out_count[names(insect_count)] <- insect_count
        out_count
      }))
      #### Plants
      plant_names <- names(unlisted_model$Z[[1]][[1]])

      plant_count <- t(sapply(plant_family, function(order) {
        out_count <- rep(0, unlisted_model$Q[1])
        names(out_count) <- seq.int(unlisted_model$Q[1])
        plant_count <- table(unlisted_model$Z[[m]][[1]][grep(order, plant_names)])
        out_count[names(plant_count)] <- plant_count
        out_count
      }))
    }
    return(list(list(insects = insect_count, plants = plant_count)))
  }
}

taxonomy_remove_empty <- function(taxonomy_collections_list) {
  lapply(taxonomy_collections_list, function(collection) {
    list(
      insects = collection$insects[which(rowSums(collection$insects != 0) > 0), ],
      plants = collection$plants[which(rowSums(collection$plants != 0) > 0), ]
    )
  })
}

taxonomy_plot <- function(data, insects_or_plants, model, stack_or_fill) {
  plots <- data %>%
    ggplot(aes(x = block, y = nb_nodes, fill = taxo)) +
    geom_bar(stat = "identity", position = stack_or_fill) +
    labs(x = "Block", y = "Number of Nodes", fill = "Taxonomy") +
    theme_minimal() +
    scale_fill_manual(values = unname(palette.colors())) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    ggtitle(paste(
      ifelse(insects_or_plants == "insects",
        "Pollinators", "Plants"
      ), "repartition (",
      ifelse(stack_or_fill == "stack", "absolute", "proportion"),
      ") in the", model, "clustering"
    )) +
    facet_wrap(~col_id, ncol = 3, scales = "free_x")

  # Arrange the plots in a grid layout
  return(gridExtra::grid.arrange(plots, newpage = TRUE))
}
```

```{r load data, echo = FALSE, include = FALSE}
# All results
iid_clustering <- readRDS("real_data/data/dore_collection_clustering_nb_run1_iid_123networks_24-05-23-21:40:42.Rds")
iid_best_partition <- extract_best_bipartite_partition(iid_clustering)
iid_unlist <- unlist(iid_best_partition)

rho_clustering <- readRDS("real_data/data/dore_collection_clustering_nb_run1_rho_123networks_25-05-23-13:58:30.Rds")
rho_best_partition <- extract_best_bipartite_partition(rho_clustering)
rho_unlist <- unlist(rho_best_partition)

pi_clustering <- readRDS("real_data/data/dore_collection_clustering_nb_run1_pi_123networks_25-05-23-17:31:25.Rds")
pi_best_partition <- extract_best_bipartite_partition(pi_clustering)
pi_unlist <- unlist(pi_best_partition)

pirho_clustering <- readRDS("real_data/data/dore_collection_clustering_nb_run1_pirho_123networks_26-05-23-19:22:55.Rds")
pirho_best_partition <- extract_best_bipartite_partition(pirho_clustering)
pirho_unlist <- unlist(pirho_best_partition)
```

## Clustering avec le modèle iid
Avec le modèle *iid* nous obtenons les `r length(iid_unlist)` collections suivantes:
```{r iid_liste_nom, echo = FALSE, results="asis"}
iid_networks_list <- lapply(seq_along(iid_unlist), function(c) {
  length <- length(iid_unlist[[c]]$net_id)
  data.frame(collection_id = rep(c, length), net_id = iid_unlist[[c]]$net_id)
})
for (df_net in iid_networks_list) {
  print(knitr::kable(df_net))
}
```

Une fois le modèle de collection appliqué nous obtenons les structures suivantes:

```{r iid_meso_plot, echo = FALSE, message=FALSE}
for (idx in seq_along(iid_unlist)) {
  cat(paste("\nPour la collection", idx, "du modèle *iid*\n"))
  print(plot(iid_unlist[[idx]], type = "meso", mixture = TRUE))
  print(iid_unlist[[idx]]$net_id)
}
```

Et voici donc les valeurs numériques pour les $\alpha$ (paramètres de connectivité).

```{r iid_alpha, echo = FALSE, results="asis"}
for (idx in seq_along(iid_unlist)) {
  cat(paste("\nPour la collection ", idx, ":\n"))
  cat(knitr::knit_print(reorder_parameters(iid_unlist[[idx]])$alpha))
}
```

### Répartition dans les clusters selon la taxonomie
```{r iid_taxonomy, echo = FALSE}
iid_taxonomy <- taxonomy_in_clusters(iid_unlist)

## Insects
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
iid_collection_taxo_insects <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(iid_collection_taxo_insects) <- col_names

for (col_id in seq_along(iid_taxonomy)) {
  iid_taxo_df <- data.frame(iid_taxonomy[[col_id]]$insects)
  colnames(iid_taxo_df) <- paste("#", seq.int(ncol(iid_taxonomy[[col_id]]$insects)), sep = "")
  iid_taxo_df <- setNames(
    iid_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(iid_taxonomy[[col_id]]$insects) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  iid_taxo_df$col_id <- rep(col_id, nrow(iid_taxo_df))

  iid_collection_taxo_insects <- rbind(iid_collection_taxo_insects, iid_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the 4 largest taxons
aggregated_taxo_insects <- aggregate(iid_collection_taxo_insects$nb_nodes, by = list(taxo = iid_collection_taxo_insects$taxo), FUN = sum)
taxos_in_other_insects <- as.character(aggregated_taxo_insects$taxo[-c(order(aggregated_taxo_insects$x, decreasing = TRUE)[1:6])])
iid_collection_taxo_insects_with_other <- iid_collection_taxo_insects %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      sum(nb_nodes[taxo %in% taxos_in_other_insects]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      "Other", as.character(taxo)
    )
  )
# Aggregating for each block
other_nb_nodes_block_insects <- aggregate(iid_collection_taxo_insects_with_other[iid_collection_taxo_insects_with_other$taxo == "Other", ]$nb_nodes, by = list(block = iid_collection_taxo_insects_with_other[iid_collection_taxo_insects_with_other$taxo == "Other", ]$block, col_id = iid_collection_taxo_insects_with_other[iid_collection_taxo_insects_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
iid_collection_taxo_insects_with_other <- rbind(iid_collection_taxo_insects_with_other[iid_collection_taxo_insects_with_other$taxo != "Other", ], other_nb_nodes_block_insects)

## Plants
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
iid_collection_taxo_plants <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(iid_collection_taxo_plants) <- col_names

for (col_id in seq_along(iid_taxonomy)) {
  iid_taxo_df <- data.frame(iid_taxonomy[[col_id]]$plants)
  colnames(iid_taxo_df) <- paste("#", seq.int(ncol(iid_taxonomy[[col_id]]$plants)), sep = "")
  iid_taxo_df <- setNames(
    iid_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(iid_taxonomy[[col_id]]$plants) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  iid_taxo_df$col_id <- rep(col_id, nrow(iid_taxo_df))

  iid_collection_taxo_plants <- rbind(iid_collection_taxo_plants, iid_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the largest taxons
aggregated_taxo_plants <- aggregate(iid_collection_taxo_plants$nb_nodes, by = list(taxo = iid_collection_taxo_plants$taxo), FUN = sum)
taxos_in_other_plants <- as.character(aggregated_taxo_plants$taxo[-c(order(aggregated_taxo_plants$x, decreasing = TRUE)[1:8])])
iid_collection_taxo_plants_with_other <- iid_collection_taxo_plants %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      sum(nb_nodes[taxo %in% taxos_in_other_plants]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      "Other", as.character(taxo)
    )
  )

# Aggregating for each block
other_nb_nodes_block_plants <- aggregate(iid_collection_taxo_plants_with_other[iid_collection_taxo_plants_with_other$taxo == "Other", ]$nb_nodes, by = list(block = iid_collection_taxo_plants_with_other[iid_collection_taxo_plants_with_other$taxo == "Other", ]$block, col_id = iid_collection_taxo_plants_with_other[iid_collection_taxo_plants_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
iid_collection_taxo_plants_with_other <- rbind(iid_collection_taxo_plants_with_other[iid_collection_taxo_plants_with_other$taxo != "Other", ], other_nb_nodes_block_plants)
```

```{r iid_plot_taxonomy_pollinators, echo = FALSE, message = FALSE,fig.cap = 'Pollinators repartition for the iid model regarding taxonomy'}
# Pollinators
taxonomy_plot(
  data = iid_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "stack", model = "iid"
)
taxonomy_plot(
  data = iid_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "fill", model = "iid"
)
```

```{r iid_plot_taxonomy_plants, echo = FALSE, message = FALSE,fig.cap = 'Plants repartition for the iid model regarding taxonomy'}
taxonomy_plot(
  data = iid_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "stack", model = "iid"
)
taxonomy_plot(
  data = iid_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "fill", model = "iid"
)
```

## Clustering avec le modèle pi
Avec le modèle *pi* nous obtenons les `r length(pi_unlist)` collections suivantes:
```{r pi_liste_nom, echo = FALSE, results="asis"}
pi_networks_list <- lapply(seq_along(pi_unlist), function(c) {
  length <- length(pi_unlist[[c]]$net_id)
  data.frame(collection_id = rep(c, length), net_id = pi_unlist[[c]]$net_id)
})
for (df_net in pi_networks_list) {
  print(knitr::kable(df_net))
}
```

Une fois le modèle de collection appliqué nous obtenons les structures suivantes:

```{r pi_meso_plot, echo = FALSE, message=FALSE}
for (idx in seq_along(pi_unlist)) {
  cat(paste("\nPour la collection", idx, "du modèle *pi*\n"))
  print(plot(pi_unlist[[idx]], type = "meso", mixture = TRUE))
  print(pi_unlist[[idx]]$net_id)
}
```

Et voici donc les valeurs numériques pour les $\alpha$ (paramètres de connectivité).

```{r pi_alpha, echo = FALSE, results="asis"}
for (idx in seq_along(pi_unlist)) {
  cat(paste("\nPour la collection ", idx, ":\n"))
  cat(knitr::knit_print(reorder_parameters(pi_unlist[[idx]])$alpha))
}
```

### Répartition dans les clusters selon la taxonomie
```{r pi_taxonomy, echo = FALSE}
pi_taxonomy <- taxonomy_in_clusters(pi_unlist)

## Insects
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
pi_collection_taxo_insects <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(pi_collection_taxo_insects) <- col_names

for (col_id in seq_along(pi_taxonomy)) {
  pi_taxo_df <- data.frame(pi_taxonomy[[col_id]]$insects)
  colnames(pi_taxo_df) <- paste("#", seq.int(ncol(pi_taxonomy[[col_id]]$insects)), sep = "")
  pi_taxo_df <- setNames(
    pi_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(pi_taxonomy[[col_id]]$insects) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  pi_taxo_df$col_id <- rep(col_id, nrow(pi_taxo_df))

  pi_collection_taxo_insects <- rbind(pi_collection_taxo_insects, pi_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the 4 largest taxons
aggregated_taxo_insects <- aggregate(pi_collection_taxo_insects$nb_nodes, by = list(taxo = pi_collection_taxo_insects$taxo), FUN = sum)
taxos_in_other_insects <- as.character(aggregated_taxo_insects$taxo[-c(order(aggregated_taxo_insects$x, decreasing = TRUE)[1:6])])
pi_collection_taxo_insects_with_other <- pi_collection_taxo_insects %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      sum(nb_nodes[taxo %in% taxos_in_other_insects]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      "Other", as.character(taxo)
    )
  )
# Aggregating for each block
other_nb_nodes_block_insects <- aggregate(pi_collection_taxo_insects_with_other[pi_collection_taxo_insects_with_other$taxo == "Other", ]$nb_nodes, by = list(block = pi_collection_taxo_insects_with_other[pi_collection_taxo_insects_with_other$taxo == "Other", ]$block, col_id = pi_collection_taxo_insects_with_other[pi_collection_taxo_insects_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
pi_collection_taxo_insects_with_other <- rbind(pi_collection_taxo_insects_with_other[pi_collection_taxo_insects_with_other$taxo != "Other", ], other_nb_nodes_block_insects)

## Plants
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
pi_collection_taxo_plants <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(pi_collection_taxo_plants) <- col_names

for (col_id in seq_along(pi_taxonomy)) {
  pi_taxo_df <- data.frame(pi_taxonomy[[col_id]]$plants)
  colnames(pi_taxo_df) <- paste("#", seq.int(ncol(pi_taxonomy[[col_id]]$plants)), sep = "")
  pi_taxo_df <- setNames(
    pi_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(pi_taxonomy[[col_id]]$plants) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  pi_taxo_df$col_id <- rep(col_id, nrow(pi_taxo_df))

  pi_collection_taxo_plants <- rbind(pi_collection_taxo_plants, pi_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the largest taxons
aggregated_taxo_plants <- aggregate(pi_collection_taxo_plants$nb_nodes, by = list(taxo = pi_collection_taxo_plants$taxo), FUN = sum)
taxos_in_other_plants <- as.character(aggregated_taxo_plants$taxo[-c(order(aggregated_taxo_plants$x, decreasing = TRUE)[1:8])])
pi_collection_taxo_plants_with_other <- pi_collection_taxo_plants %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      sum(nb_nodes[taxo %in% taxos_in_other_plants]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      "Other", as.character(taxo)
    )
  )

# Aggregating for each block
other_nb_nodes_block_plants <- aggregate(pi_collection_taxo_plants_with_other[pi_collection_taxo_plants_with_other$taxo == "Other", ]$nb_nodes, by = list(block = pi_collection_taxo_plants_with_other[pi_collection_taxo_plants_with_other$taxo == "Other", ]$block, col_id = pi_collection_taxo_plants_with_other[pi_collection_taxo_plants_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
pi_collection_taxo_plants_with_other <- rbind(pi_collection_taxo_plants_with_other[pi_collection_taxo_plants_with_other$taxo != "Other", ], other_nb_nodes_block_plants)
```

```{r pi_plot_taxonomy_pollinators, echo = FALSE, message = FALSE,fig.cap = 'Pollinators repartition for the pi model regarding taxonomy'}
# Pollinators
taxonomy_plot(
  data = pi_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "stack", model = "pi"
)
taxonomy_plot(
  data = pi_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "fill", model = "pi"
)
```

```{r pi_plot_taxonomy_plants, echo = FALSE, message = FALSE,fig.cap = 'Plants repartition for the pi model regarding taxonomy'}
taxonomy_plot(
  data = pi_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "stack", model = "pi"
)
taxonomy_plot(
  data = pi_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "fill", model = "pi"
)
```

## Clustering avec le modèle rho
Avec le modèle *rho* nous obtenons les `r length(rho_unlist)` collections suivantes:
```{r rho_liste_nom, echo = FALSE, results="asis"}
print(knitr::kable(rho_unlist$net_id))
```

Une fois le modèle de collection appliqué nous obtenons les structures suivantes:

```{r rho_meso_plot, echo = FALSE, message=FALSE}
cat(paste("\nPour la collection 1 du modèle *rho*\n"))
print(plot(rho_unlist, type = "meso", mixture = TRUE))
print(rho_unlist$net_id)
```

Et voici donc les valeurs numériques pour les $\alpha$ (paramètres de connectivité).

```{r rho_alpha, echo = FALSE, results="asis"}
cat(paste("\nPour la collection 1:\n"))
cat(knitr::knit_print(reorder_parameters(rho_unlist)$alpha))
```

### Répartition dans les clusters selon la taxonomie
```{r rho_taxonomy, echo = FALSE}
rho_taxonomy <- taxonomy_in_clusters(rho_unlist)

## Insects
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
rho_collection_taxo_insects <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(rho_collection_taxo_insects) <- col_names

for (col_id in seq_along(rho_taxonomy)) {
  rho_taxo_df <- data.frame(rho_taxonomy[[col_id]]$insects)
  colnames(rho_taxo_df) <- paste("#", seq.int(ncol(rho_taxonomy[[col_id]]$insects)), sep = "")
  rho_taxo_df <- setNames(
    rho_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(rho_taxonomy[[col_id]]$insects) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  rho_taxo_df$col_id <- rep(col_id, nrow(rho_taxo_df))

  rho_collection_taxo_insects <- rbind(rho_collection_taxo_insects, rho_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the 4 largest taxons
aggregated_taxo_insects <- aggregate(rho_collection_taxo_insects$nb_nodes, by = list(taxo = rho_collection_taxo_insects$taxo), FUN = sum)
taxos_in_other_insects <- as.character(aggregated_taxo_insects$taxo[-c(order(aggregated_taxo_insects$x, decreasing = TRUE)[1:6])])
rho_collection_taxo_insects_with_other <- rho_collection_taxo_insects %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      sum(nb_nodes[taxo %in% taxos_in_other_insects]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      "Other", as.character(taxo)
    )
  )
# Aggregating for each block
other_nb_nodes_block_insects <- aggregate(rho_collection_taxo_insects_with_other[rho_collection_taxo_insects_with_other$taxo == "Other", ]$nb_nodes, by = list(block = rho_collection_taxo_insects_with_other[rho_collection_taxo_insects_with_other$taxo == "Other", ]$block, col_id = rho_collection_taxo_insects_with_other[rho_collection_taxo_insects_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
rho_collection_taxo_insects_with_other <- rbind(rho_collection_taxo_insects_with_other[rho_collection_taxo_insects_with_other$taxo != "Other", ], other_nb_nodes_block_insects)

## Plants
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
rho_collection_taxo_plants <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(rho_collection_taxo_plants) <- col_names

for (col_id in seq_along(rho_taxonomy)) {
  rho_taxo_df <- data.frame(rho_taxonomy[[col_id]]$plants)
  colnames(rho_taxo_df) <- paste("#", seq.int(ncol(rho_taxonomy[[col_id]]$plants)), sep = "")
  rho_taxo_df <- setNames(
    rho_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(rho_taxonomy[[col_id]]$plants) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  rho_taxo_df$col_id <- rep(col_id, nrow(rho_taxo_df))

  rho_collection_taxo_plants <- rbind(rho_collection_taxo_plants, rho_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the largest taxons
aggregated_taxo_plants <- aggregate(rho_collection_taxo_plants$nb_nodes, by = list(taxo = rho_collection_taxo_plants$taxo), FUN = sum)
taxos_in_other_plants <- as.character(aggregated_taxo_plants$taxo[-c(order(aggregated_taxo_plants$x, decreasing = TRUE)[1:8])])
rho_collection_taxo_plants_with_other <- rho_collection_taxo_plants %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      sum(nb_nodes[taxo %in% taxos_in_other_plants]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      "Other", as.character(taxo)
    )
  )

# Aggregating for each block
other_nb_nodes_block_plants <- aggregate(rho_collection_taxo_plants_with_other[rho_collection_taxo_plants_with_other$taxo == "Other", ]$nb_nodes, by = list(block = rho_collection_taxo_plants_with_other[rho_collection_taxo_plants_with_other$taxo == "Other", ]$block, col_id = rho_collection_taxo_plants_with_other[rho_collection_taxo_plants_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
rho_collection_taxo_plants_with_other <- rbind(rho_collection_taxo_plants_with_other[rho_collection_taxo_plants_with_other$taxo != "Other", ], other_nb_nodes_block_plants)
```

```{r rho_plot_taxonomy_pollinators, echo = FALSE, message = FALSE,fig.cap = 'Pollinators repartition for the rho model regarding taxonomy'}
# Pollinators
taxonomy_plot(
  data = rho_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "stack", model = "rho"
)
taxonomy_plot(
  data = rho_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "fill", model = "rho"
)
```

```{r rho_plot_taxonomy_plants, echo = FALSE, message = FALSE,fig.cap = 'Plants repartition for the rho model regarding taxonomy'}
taxonomy_plot(
  data = rho_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "stack", model = "rho"
)
taxonomy_plot(
  data = rho_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "fill", model = "rho"
)
```

## Clustering avec le modèle pirho
Avec le modèle *pirho* nous obtenons les `r length(pirho_unlist)` collections suivantes:
```{r pirho_liste_nom, echo = FALSE, results="asis"}
pirho_networks_list <- lapply(seq_along(pirho_unlist), function(c) {
  length <- length(pirho_unlist[[c]]$net_id)
  data.frame(collection_id = rep(c, length), net_id = pirho_unlist[[c]]$net_id)
})
for (df_net in pirho_networks_list) {
  print(knitr::kable(df_net))
}
```

Une fois le modèle de collection appliqué nous obtenons les structures suivantes:

```{r pirho_meso_plot, echo = FALSE, message=FALSE}
for (idx in seq_along(pirho_unlist)) {
  cat(paste("\nPour la collection", idx, "du modèle *pirho*\n"))
  print(plot(pirho_unlist[[idx]], type = "meso", mixture = TRUE))
  print(pirho_unlist[[idx]]$net_id)
}
```

Et voici donc les valeurs numériques pour les $\alpha$ (paramètres de connectivité).

```{r pirho_alpha, echo = FALSE, results="asis"}
for (idx in seq_along(pirho_unlist)) {
  cat(paste("\nPour la collection ", idx, ":\n"))
  cat(knitr::knit_print(reorder_parameters(pirho_unlist[[idx]])$alpha))
}
```

### Répartition dans les clusters selon la taxonomie
```{r pirho_taxonomy, echo = FALSE}
pirho_taxonomy <- taxonomy_in_clusters(pirho_unlist)

## Insects
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
pirho_collection_taxo_insects <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(pirho_collection_taxo_insects) <- col_names

for (col_id in seq_along(pirho_taxonomy)) {
  pirho_taxo_df <- data.frame(pirho_taxonomy[[col_id]]$insects)
  colnames(pirho_taxo_df) <- paste("#", seq.int(ncol(pirho_taxonomy[[col_id]]$insects)), sep = "")
  pirho_taxo_df <- setNames(
    pirho_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(pirho_taxonomy[[col_id]]$insects) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  pirho_taxo_df$col_id <- rep(col_id, nrow(pirho_taxo_df))

  pirho_collection_taxo_insects <- rbind(pirho_collection_taxo_insects, pirho_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the 4 largest taxons
aggregated_taxo_insects <- aggregate(pirho_collection_taxo_insects$nb_nodes, by = list(taxo = pirho_collection_taxo_insects$taxo), FUN = sum)
taxos_in_other_insects <- as.character(aggregated_taxo_insects$taxo[-c(order(aggregated_taxo_insects$x, decreasing = TRUE)[1:6])])
pirho_collection_taxo_insects_with_other <- pirho_collection_taxo_insects %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      sum(nb_nodes[taxo %in% taxos_in_other_insects]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_insects,
      "Other", as.character(taxo)
    )
  )
# Aggregating for each block
other_nb_nodes_block_insects <- aggregate(pirho_collection_taxo_insects_with_other[pirho_collection_taxo_insects_with_other$taxo == "Other", ]$nb_nodes, by = list(block = pirho_collection_taxo_insects_with_other[pirho_collection_taxo_insects_with_other$taxo == "Other", ]$block, col_id = pirho_collection_taxo_insects_with_other[pirho_collection_taxo_insects_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
pirho_collection_taxo_insects_with_other <- rbind(pirho_collection_taxo_insects_with_other[pirho_collection_taxo_insects_with_other$taxo != "Other", ], other_nb_nodes_block_insects)

## Plants
# Prepare data for ggplot
col_names <- c("taxo", "block", "nb_nodes", "col_id")
pirho_collection_taxo_plants <- data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(pirho_collection_taxo_plants) <- col_names

for (col_id in seq_along(pirho_taxonomy)) {
  pirho_taxo_df <- data.frame(pirho_taxonomy[[col_id]]$plants)
  colnames(pirho_taxo_df) <- paste("#", seq.int(ncol(pirho_taxonomy[[col_id]]$plants)), sep = "")
  pirho_taxo_df <- setNames(
    pirho_taxo_df %>%
      tibble::rownames_to_column(var = "taxo") %>%
      tidyr::pivot_longer(cols = 2:(ncol(pirho_taxonomy[[col_id]]$plants) + 1)),
    c("taxo", "block", "nb_nodes")
  )

  pirho_taxo_df$col_id <- rep(col_id, nrow(pirho_taxo_df))

  pirho_collection_taxo_plants <- rbind(pirho_collection_taxo_plants, pirho_taxo_df %>%
    mutate(block = factor(block), taxo = factor(taxo), col_id = factor(col_id)))
}

# Find the largest taxons
aggregated_taxo_plants <- aggregate(pirho_collection_taxo_plants$nb_nodes, by = list(taxo = pirho_collection_taxo_plants$taxo), FUN = sum)
taxos_in_other_plants <- as.character(aggregated_taxo_plants$taxo[-c(order(aggregated_taxo_plants$x, decreasing = TRUE)[1:8])])
pirho_collection_taxo_plants_with_other <- pirho_collection_taxo_plants %>%
  group_by(col_id, taxo, block) %>%
  summarize(
    nb_nodes = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      sum(nb_nodes[taxo %in% taxos_in_other_plants]), nb_nodes
    ),
    taxo = ifelse(as.character(taxo) %in% taxos_in_other_plants,
      "Other", as.character(taxo)
    )
  )

# Aggregating for each block
other_nb_nodes_block_plants <- aggregate(pirho_collection_taxo_plants_with_other[pirho_collection_taxo_plants_with_other$taxo == "Other", ]$nb_nodes, by = list(block = pirho_collection_taxo_plants_with_other[pirho_collection_taxo_plants_with_other$taxo == "Other", ]$block, col_id = pirho_collection_taxo_plants_with_other[pirho_collection_taxo_plants_with_other$taxo == "Other", ]$col_id), FUN = sum) %>%
  mutate(taxo = factor("Other"), nb_nodes = x, x = NULL)
# Removing the different other and remerging with the aggregated
pirho_collection_taxo_plants_with_other <- rbind(pirho_collection_taxo_plants_with_other[pirho_collection_taxo_plants_with_other$taxo != "Other", ], other_nb_nodes_block_plants)
```

```{r pirho_plot_taxonomy_pollinators, echo = FALSE, message = FALSE,fig.cap = 'Pollinators repartition for the pirho model regarding taxonomy'}
# Pollinators
taxonomy_plot(
  data = pirho_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "stack", model = "pirho"
)
taxonomy_plot(
  data = pirho_collection_taxo_insects_with_other,
  insects_or_plants = "insects",
  stack_or_fill = "fill", model = "pirho"
)
```

```{r pirho_plot_taxonomy_plants, echo = FALSE, message = FALSE,fig.cap = 'Plants repartition for the pirho model regarding taxonomy'}
taxonomy_plot(
  data = pirho_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "stack", model = "pirho"
)
taxonomy_plot(
  data = pirho_collection_taxo_plants_with_other,
  insects_or_plants = "plants",
  stack_or_fill = "fill", model = "pirho"
)
```
