---
title: "Analyzing the capacity of the colBiSBM to recover structure for missing data from other networks"
output:
    html_document:
        toc: true
        theme: journal
    pdf_document:
        keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r required_libs, echo = FALSE, include=FALSE}
require("ggplot2")
require("tidyverse")
```

```{r import_data, echo=FALSE, include=FALSE}
NA_robustness_raw <- readRDS("simulation/data/NA_robustness_results-alpha_0.7, 0.4, 0.3, 0.4, 0.2, 0.05, 0.3, 0.05, 0.05-reps-10-14-06-23_17-41.Rds")
NA_robustness_df <- NA_robustness_raw %>%
  mutate(
    auc_diff = auc_colBiSBM - auc_LBM,
    ari_row_diff = NA_net_ari_row - LBM_ari_row,
    ari_col_diff = NA_net_ari_col - LBM_ari_col
  ) %>%
  group_by(prop_NAs, model) %>%
  summarise(
    mean_auc_diff = mean(auc_diff),
    sd_auc_diff = sd(auc_diff),
    mean_ari_row_diff = mean(ari_row_diff),
    sd_ari_row_diff = sd(ari_row_diff),
    mean_ari_col_diff = mean(ari_col_diff),
    sd_ari_col_diff = sd(ari_col_diff),
    mean_LBM_ari_row = mean(LBM_ari_row),
    sd_LBM_ari_row = sd(LBM_ari_row),
    mean_LBM_ari_col = mean(LBM_ari_col),
    sd_LBM_ari_col = sd(LBM_ari_col),
    mean_NA_net_ari_row = mean(NA_net_ari_row),
    sd_NA_net_ari_row = sd(NA_net_ari_row),
    mean_NA_net_ari_col = mean(NA_net_ari_col),
    sd_NA_net_ari_col = sd(NA_net_ari_col),
    mean_elapsed_secs = mean(elapsed_secs),
    sd_elapsed_secs = sd(elapsed_secs)
  ) %>%
  ungroup()
```

```{r useful_function, echo = FALSE}
write_matex2 <- function(x) {
  if (!is.matrix(x)) {
    x <- matrix(x)
  }
  begin <- "\\begin{bmatrix}"
  end <- "\\end{bmatrix}"
  X <-
    apply(x, 1, function(x) {
      paste(
        paste(x, collapse = "&"),
        "\\\\"
      )
    })
  paste(c(begin, X, end), collapse = "")
}
```

# Simulation context

The idea is to benchmark the capacity of the models when NAs are in the data.

To do this, we chose the below structure:

```{r simulation_parameters, echo = FALSE}
eps <- 0.05

M <- 3

# Defining parameters
nr <- 100
nc <- 150
pir <- c(0.5, 0.3, 0.2)
pic <- c(0.5, 0.3, 0.2)
alpha <- matrix(c(
  0.6, eps, eps,
  eps, 0.4, eps,
  eps, eps, 0.7
), byrow = TRUE, nrow = length(pir), ncol = length(pic))
```

$$M = `r M`, n_r = `r nr`, n_c = `r nc` \\ \alpha = `r write_matex2(alpha)`
    \\ \pi = `r write_matex2(pir)` \rho = `r write_matex2(pic)`$$

With $M$ the number of networks, $n_r$ the number of nodes in row of the incidence
matrix, $n_c$ the number of nodes in column, $\alpha$ the connectivity
parameters between the row and column clusters. $\pi$ and $\rho$ are
the proportion of nodes in the row and columns clusters.

And set some randomly chosen interactions to NA. The below plots will show the
different quality indicators in function of proportion of NAs in the first of
the 3 networks.

# AUC in function of the proportion of NAs

```{r auc_plots, echo = FALSE}
auc_plot <- NA_robustness_df %>% ggplot() +
  geom_line(aes(x = prop_NAs, y = mean_auc_diff, color = model)) +
  geom_point(aes(x = prop_NAs, y = mean_auc_diff, color = model)) +
  geom_ribbon(aes(ymin = mean_auc_diff - sd_auc_diff, ymax = mean_auc_diff + sd_auc_diff, x = prop_NAs, fill = model), alpha = 0.25) +
  xlab("NA proportion") +
  ylab("AUC difference (colBiSBM - LBM)") +
  scale_x_continuous(breaks = seq(0, 0.9, 0.1))


auc_plot
```

<!-- FAUX
    At $0$, there is no NAs in the $1^{st}$ network so the information is completely
retrieved.

At $0.1$, the AUC is degraded and falls around 0.85 for the *iid, pi and rho*
models. The *pirho* model seems to perform slightly better, but it's confidence interval
intersects the others.

For the next values, two "behaviours" can be observed :
- the *iid* decreases and seems to stabilize until 0.8.
- the *pi, rho and pirho* seem to remain stable until 0.7, from 0.8 they go down.
But the *pirho* performs overall the best, staying around 0.86 and with $90\%$
NAs in ther first network it gives an AUC of 0.83 !

The *pi* and *rho* models perform the same (their confidence intervals overlap),
and they give an AUC of around 0.83 between $20\%$ to $80\%$ of NAs. The *rho*
ends at 0.8 AUC, where the *pi* ends at 0.76.

The 4 models maintain an AUC over 0.75 for $90%$ -->

# ARI in function of the proportion of NAs

```{r ARI_row_plot, echo = FALSE, fig.cap="Difference of ARI for the row clusterings"}
ari_row_plot <- NA_robustness_df %>% ggplot() +
  ylim(-1, 1) +
  geom_line(aes(x = prop_NAs, y = mean_ari_row_diff, color = model)) +
  geom_point(aes(x = prop_NAs, y = mean_ari_row_diff, color = model)) +
  geom_ribbon(aes(ymin = mean_ari_row_diff - sd_ari_row_diff, ymax = mean_ari_row_diff + sd_ari_row_diff, x = prop_NAs, fill = model), alpha = 0.25) +
  xlab("NA proportion") +
  ylab("ARI difference") +
  ggtitle("ARI on the row clustering, difference (colBiSBM - LBM)") +
  scale_x_continuous(breaks = seq(0, 0.9, 0.1))

ari_row_plot
```

```{r ARI_col_plot, echo = FALSE, fig.cap="Difference of ARI for the columns clusterings"}
ari_col_plot <- NA_robustness_df %>% ggplot() +
  ylim(-1, 1) +
  geom_line(aes(x = prop_NAs, y = mean_ari_col_diff, color = model)) +
  geom_point(aes(x = prop_NAs, y = mean_ari_col_diff, color = model)) +
  geom_ribbon(aes(ymin = mean_ari_col_diff - sd_ari_col_diff, ymax = mean_ari_col_diff + sd_ari_col_diff, x = prop_NAs, fill = model), alpha = 0.25) +
  xlab("NA proportion") +
  ylab("ARI difference") +
  ggtitle("ARI on the column clustering, difference (colBiSBM - LBM)") +
  scale_x_continuous(breaks = seq(0, 0.9, 0.1))

ari_col_plot
```
